---
title: "Data Import"
author: "Kelly Wang"
date: "9/17/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

## Load in a dataset

```{r}
##reads in a dataset
litters_data=read_csv(file ="./data/FAS_litters.csv")
litters_data=janitor::clean_names(litters_data)
```

#load in pups data

```{r}
pups_data=read_csv(file="./data/FAS_pups.csv")
pups_data=janitor::clean_names(litters_data)
```
##play with column types
litters_data = read_csv(
  file = "./data/FAS_litters.csv",
  col_types = cols(
    Group = col_character(),
    `Litter Number` = col_character(),
    `GD0 weight` = col_double(),
    `GD18 weight` = col_double(),
    `GD of Birth` = col_integer(),
    `Pups born alive` = col_integer(),
    `Pups dead @ birth` = col_integer(),
    `Pups survive` = col_integer()
  )
)
tail(litters_data)

##read in an excel file
```{r}
library(readxl)

mlb11_data=read_excel(path="./data/mlb11.xlsx")
#only if you just want a subset of the excel sheet
mlb11_data_subset=
  read_excel(
    path="./data/mlb11.xlsx",
    range="A1:A77"
  )
```

##read in SAS...
```{r}
pulse_data=haven::read_sas("./data/public_pulse_data.sas7bdat")
```

#exporting dataset
```{r}
write_csv(mlb11_data_subset, path = "./data/mlb11_subset.csv")
```