---
title: "Data Import"
author: "Kelly Wang"
date: "9/19/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

## Import dataset

```{r}
##reads in a dataset
litters_data=read_csv(file ="./data/FAS_litters.csv")
litters_data=janitor::clean_names(litters_data)
pups_data=read_csv(file="./data/FAS_pups.csv")
pups_data=janitor::clean_names(litters_data)
```

R doesn't know what group is. however we can say that we know that inside the litters data set there is a variable group. R only knows about the data frames and what is assigned. There is a difference here that R knows about directly and R knows if we tell them where to look, which is inside the data from

## Selecting!! 
```{r}
select(litters_data,group,litter_number, gd0_weight, starts_with("pups"))

select(litters_data, litter_number, group, gd0_weight)

#if you just want everything but one variable. the minus sign indicates subtracting the variable
select(litters_data, -group)

#to rearrange
select(litters_data, gd18_weight, everything())

#keep every column between gd0_weight and pups_born_alive
select(litters_data, litter_number, gd0_weight:pups_born_alive)

#how to rename variable names
select(litters_data, GROUP = group, litter_number)
```
first thing is data and then after that are the columns you want 
note that starts_with is a helper function. it allows us to chose a key word that starts with pups 

note the select does put variables in order

starts_with(), ends_with() and contains()

## Filtering
```{r}
filter(litters_data, group == "Mod8")
## == when this is true, keep this row

filter(litters_data, gd_of_birth == 20)

filter(litters_data, gd_of_birth < 20)

filter(litters_data, gd_of_birth >= 20)

filter(litters_data, pups_born_alive < 6)

#either in this group or other
filter(litters_data,group %in% c("Con7", "Con8"))

#!(pups_survive == 4) means not 

#multiple conditions

filter(litters_data, pups_born_alive >= 4, pups_born_alive <=6)

#a+b>7
filter(litters_data, gd0_weight+gd18_weight>70)

#what if there are missing values?
# don't do this filter(litters_data, !is.na(gd0_weight))

#recommends this:
drop_na(litters_data)
```


